Rectangulotopes

### 2024-04-04 draft/rectangulotopes.tex
proposition change thm 1

Among those, the \defn{associahedron} is a classical, ubiquitous polytope, first defined by Tamari~\cite{T51}, then studied by Stasheff in a topological context~\cite{S63}, that has since then found applications in various areas of mathematics, including topology, operads, cluster algebras, combinatorial Hopf algebras, and physics~\cite{MR4675114}.
It was first defined as a combinatorial object, before various families of geometric realizations were shown to exist~\cite{MR1022776,MR1941227,MR2108555,MR3437894}.
The face lattice of the $(n-1)$-dimensional associahedron is the reverse inclusion poset of nonintersecting diagonals of a convex $(n+3)$-gon.
In particular, it skeleton is the \defn{flip graph} on \defn{triangulations} of the $(n+3)$-gon, or equivalently -- via a standard Catalan bijection -- the \defn{rotation graph} on \defn{binary trees} with $n$ internal nodes, the structure of which has been the subject of many investigations, with applications in computer science~\cite{MR928904,MR3197650}.
In 2004, Loday published the following elegant realization of the associahedron~\cite{MR2108555}, giving a recipe for computing the coordinates of a vertex given the corresponding binary tree.

@@

@@

By definition, any quotientope is a generalized (or deformed) permutahedron, and can thus be written as
  \[
  \Bigset{x \in \R^n}{\sum_{i \in S} x_i \le f(S)\ \forall \varnothing \ne S \subseteq [n] \text { and } \sum_{i \in [n]} x_i = f([n])},
  \]
where~$f$ is a \defn{submodular} set function $f:2^{[n]}\mapsto \R$ (\ie such that~$f(X \cup Y) + f(X \cap Y) \le f(X) + f(Y)$ for any~$X, Y \subseteq [n]$).

>>>>

Among those, the \defn{associahedron} is a classical, ubiquitous polytope, first defined by Tamari~\cite{T51}, then studied by Stasheff in a topological context~\cite{S63}, that has since then found applications in various areas of mathematics, including topology, operads, cluster algebras, combinatorial Hopf algebras, and physics~\cite{MR4675114}.
It was first defined as a combinatorial object, before various families of geometric realizations were shown to exist~\cite{MR1022776,MR1941227,MR2108555,MR3437894}.
The face lattice of the $(n-1)$-dimensional associahedron is the reverse inclusion poset of nonintersecting diagonals of a convex $(n+3)$-gon.
In particular, it skeleton is the \defn{flip graph} on \defn{triangulations} of the $(n+3)$-gon, or equivalently -- via a standard Catalan bijection -- the \defn{rotation graph} on \defn{binary trees} with $n$ internal nodes, the structure of which has been the subject of many investigations, with applications in computer science~\cite{MR928904,MR3197650}.

\Vincent{I would change this theorem...}
In 2004, Loday published the following elegant realization of the associahedron~\cite{MR2108555}, giving a recipe for computing the coordinates of a vertex given the corresponding binary tree.

@@

\Vincent{... as follows}

\begin{theorem}
  \label{thm:loday}
  The associahedron is realized by the polytope~$\Asso (n)$ in~$\R^n$ defined equivalently as
  \begin{enumerate}[(i)]
  \item the convex hull of the points~$\sum_{i\in [n]} \ell^T_i\cdot r^T_i \cdot \mathbf{e}_i$ for all binary trees~$T$ with $n$ internal nodes, where $\ell^T_i$ and $r^T_i$ denote the number of leaves in the left and right subtrees of $i$ in $T$,
  \item the intersection of the hyperplane defined by~$\sum_{i \in [n]} x_i = \binom{n+1}{2}$ with the halfspaces defined by~$\sum_{i \in S} x_i \le \#\set{I \text{ interval of } [n]}{S \cap I \ne \varnothing}$.
  \end{enumerate}
\end{theorem}

@@

By definition, any quotientope is a deformed permutahedron, and can thus be written as
  \[
  \Bigset{x \in \R^n}{\sum_{i \in S} x_i \le f(S)\ \forall \varnothing \ne S \subseteq [n] \text { and } \sum_{i \in [n]} x_i = f([n])},
  \]
where~$f$ is a \defn{submodular} set function $f:2^{[n]}\mapsto \R$ (\ie such that~$f(X \cup Y) + f(X \cap Y) \le f(X) + f(Y)$ for any~$X, Y \subseteq [n]$).
Recall that the submodular function of a Minkowski sum of deformed permutahedra is just the sum of the submodular functions of the Minkowski summands.
